---
title: "index"
author: "Julia Mason"
date: "11/30/2017"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
library (beyonce)

```

## MPG dataset

Use data() to view available R datasets. Use ?mpg to read more about the specific dataset/metadata. 

Use backticks around words to put them in "code font" E.g. `geom_scatter`

### Scatterplot
```{r mpg}
# mpg

g <- ggplot (data = mpg,
             aes (x = displ, y = hwy)) 
g + geom_point()

```

#### Color! Going DEEPER INTO HEADINGS
```{r color scatter}
g + geom_point (aes (color = class))
```

#### Use shape for a b&w figure
Changed `color` to `shape`

```{r b&w}
g + geom_point (aes (shape = class))
```
```{r color and size}
g + geom_point (aes (color = fl, size = class))
```
#### 5.5 Exercises:
Scatter hwy vs cty, size represents class and color represents fuel type

```{r scatter exercise}
ggplot (mpg, aes (x = cty, y = hwy, size = class, col = fl)) +
  geom_point()
```

1: color = "blue" should be outside of aesthetic. 
```{r 5.5 1}

ggplot(data = mpg) + 
      geom_point(aes(x = displ, y = hwy, color = "blue")) +
  ggtitle ("Wrong")

ggplot(data = mpg) + 
      geom_point(aes(x = displ, y = hwy), color = "blue") +
  ggtitle ("Right")
```


2: plot hw vs disply, color indicates city
```{r 5.5 2}
ggplot (mpg, aes (x = displ, y = hwy, col = cty)) +
  geom_point()
```
3: now use shape for cty. Doesn't work bc cty is a continuous variable--can't have a shape mapping. 
```{r 5.5 3, eval = FALSE}
ggplot (mpg, aes (x = displ, y = hwy, shape = cty)) +
  geom_point()
```

4: map aesthetic to something other than variable, e.g displ < 5. Yields two colors reflecting true/false
```{r 5.5 4}

ggplot (mpg, aes (x = displ, y = hwy, col = displ < 5)) +
  geom_point()

```

## Geoms

```{r geom_examples}

ggplot (mpg, aes (x = drv, y = hwy)) +
  geom_jitter()

ggplot (mpg, aes (x = drv, y = hwy)) +
  geom_violin()


```

## Geoms on geoms
```{r geomception}

ggplot (mpg, aes (x = displ, y = hwy)) +
  geom_point () +
  geom_smooth ()

```
## Customize
```{r customize}
ggplot (mpg, aes (x = displ, y = hwy)) +
  geom_point () +
  geom_smooth () +
  labs (title = "Rltp btw engine size and hwy mpg",
       x = "Engine displacement (l)", 
       y = "Hwy mpg") +
  theme_minimal () +
  theme (text = element_text(size = 16))
```

### A CHALLENGE (5.9)
Create a beautiful graph
```{r beautify challenge}


ggplot (mpg, aes (x = class, y = hwy, fill = class)) +
  geom_violin(size = 0) + # removes outline
  theme_bw() +
  scale_fill_manual (values = beyonce_palette(74)) +
  labs (fill = "Car type") +
  theme (axis.text = element_text (angle = 45, hjust = 1))
  

```
